### 2018-02-27 E1/mixModelE1data.Rmd
mix model initial pilot

---
title: "mixModelE1data"
author: "Alex Holcombe"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,  comment = NA)
# http://htmlpreview.github.io/?https://github.com/alexholcombe/MixtureModelRSVP/blob/master/compareMATLABtoR.html
```

Load data

```{r load data, echo=FALSE, message=FALSE}

#Import raw data
dataPath<- file.path("Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E1_rawDataFromMAT.rda") )

#tidy data
library(dplyr)
df<- data
#to work with dplyr, can't have array field like letterSeq
df$letterSeq<- NULL

numItemsInStream<- length( data$letterSeq[1,] )  
minSPE<- -17; maxSPE<- 17

df<-df %>% rename(stream = target, orientation = condition )
df <- df %>% mutate( stream =ifelse(stream==1, "Left","Right") )
#mutate condition to Orientation
df <- df %>% mutate( orientation =ifelse(orientation==1, "Canonical","Inverted") )

```

Fit mixture model with R (if fitting not previously done and cached). 
```{r, echo=FALSE, cache=TRUE}
#Setting cache = TRUE in hopes won't have to recalculate estimates "if cached results exist and this code chunk has not been changed since last run" 
library(mixRSVP)

#Check whether already have parameter estimates or instead need to do it
calculate<-FALSE
if (!exists("estimates")) { 
  calculate<-TRUE 
}

condtnVariableNames <- c("subject","stream", "orientation")

if (calculate) {
  estimates<- df %>%   filter(subject=="AA") %>%
    group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
    do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )
  estimates<- estimates %>% rename(efficacy = p1, latency = p2, precision = p3)

}
head(estimates)
```

Save the param estimates to a file
```{Save param estimates, echo=FALSE}

write.csv(estimates, file = "Data/backwards2E1_paramEstimates")

```

Let's inspect histogram plots and fits. Note that in every case, the negative log likelihood is better (smaller) for R, showing that the R code provides a better fit than the MATLAB code.

* yellow = guessing component 
* light blue = Gaussian component
* green = sum of the guessing and Gaussian components. In other words, the histogram heights predicted by the model
* dark blue = continuous Gaussian. This helps get a sense of the effect of discretising the Gaussian. For instance, it's possible (especially using Pat's method, it seems), for the Gaussian peak to fly high above the bars and still fit the discrete bins (or bin centers, in Pat's method), suggesting an undesirably high estimates of the efficacy (likely accompanied by an undesirably low precision)

```{r, echo=FALSE, message=FALSE, fig.height=360, fig.width=10}
#want fig.height of 10 per subject

source( file.path(mixModelingPath,"calcCurvesDataframes.R") )

#create R curves
df<- df # %>%  dplyr::filter(subject <="BB") #dplyr::filter(subject <= "BD" & subject >="AP")
#dplyr::filter(subject=="AA",orientation=="Canonical")
source( file.path(mixModelingPath,"histogramPlotting.R") ) #for calcFitDataframes
source( file.path(mixModelingPath,"theme_apa.R") ) #for calcFitDataframes

#Add R parameter estimates to dataframe
df<- merge(df,estimates)

curvesR<- df %>% group_by_at(.vars = condtnVariableNames) %>% 
  do(calcCurvesDataframes(.,minSPE,maxSPE,numItemsInStream))

#Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
#Calc the number of observations for each condition, because gaussianScaledforData needs to know.
dfGroups<- df %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
#add nPerCond back to parameter estimates
estimates<- merge(estimates,dfGroups)
grain<-.05
gaussianFineR<- estimates %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussianScaledFromDataframe(.,minSPE,maxSPE,grain) )

#create MATLAB curves
#need to rename parameters so that calcCurvesDataframes can find them
estimates_M<- estimates_MATLAB %>% rename(efficacy=efficacy_M,latency=latency_M,precision=precision_M)
#create temporary dataframe with data plus MATLAB estimates to generate curves
dataStrippedOfEstimates<- select(df,-efficacy,-latency,-precision, -val)
dM<-merge( dataStrippedOfEstimates, estimates_M )
dM<- dM %>% rename( val = val_M ) #has to be called val so calcFitDataframes uses it rather than using R to calculate its own
curvesMATLAB<- dM %>% group_by_at(.vars = condtnVariableNames) %>% 
            do(calcCurvesDataframes(.,minSPE,maxSPE,numItemsInStream))

#Calculate fine-grained Gaussian for MATLAB
#Calc the number of observations for each condition, because gaussianScaledforData needs to know.
dNum<- dataStrippedOfEstimates %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
#add nPerCond to MATLAB estimates, which will then be merged with raw data
estimates_M<- merge(estimates_M,dNum)
#merge MATLAB estimates with raw data
dM<-merge( dataStrippedOfEstimates, estimates_M )
gaussianFineM<- estimates_M %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussianScaledFromDataframe(.,minSPE,maxSPE,grain) )

#Concatenate R and MATLAB into single dataframe
curvesR$lang<-"R"; curvesMATLAB$lang<-"MATLAB"
curves_R_MATLAB <- rbind( data.frame(curvesR), data.frame(curvesMATLAB) )
#To facilitate plotting,
#make dataframe containing the raw data twice, one for R plots column, one for MATLAB plots column (data identical)
dfBoth<-rbind(   df %>% mutate(lang="R"), df %>% mutate(lang="MATLAB")  )

#Concatenate fine-grained gaussian for R and MATLAB into single dataframe, so can plot both with ggplot
gaussianFineR$lang<-"R"; gaussianFineM$lang<-"MATLAB"; 
#Get the col names to match by only using a few columns
gaussianFineR<- gaussianFineR %>% select(-val,-warnings)
gaussianFineM<- gaussianFineM %>% select(-val_M)
gaussFine_R_M <- rbind( data.frame(gaussianFineR), data.frame(gaussianFineM) )

#PLOT EVERYTHING
g=ggplot(dfBoth, aes(x=SPE)) + facet_grid(subject+stream+orientation~lang,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
#g<-g+ theme(line=element_blank(), panel.border = element_blank())
sz=.8
#Plot the underlying Gaussian , not just the discretized Gaussian
g<-g + geom_line(data=gaussFine_R_M,aes(x=x,y=gaussianFreq),color="darkblue",size=1.2)

g<-g+ geom_point(data=curves_R_MATLAB,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz*2.5)
g<-g+ geom_line(data=curves_R_MATLAB,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
g<-g+ geom_line(data=curves_R_MATLAB,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

numGroups<- length(table(dfBoth$orientation,dfBoth$subject,dfBoth$stream))
fontSz = 400/numGroups
g<-g + geom_text(data=curves_R_MATLAB,aes(x=-9,y=32, label = paste("-logLik==", round(val,1), sep = "")),  parse = TRUE,size=fontSz) +
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=28, label = paste("plain(e)==", round(efficacy,2), sep = "")),  parse = TRUE,size=fontSz) +
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=25, label = paste("mu==", round(latency,2), sep = "")),  parse = TRUE,size=fontSz)+
  geom_text(data=curves_R_MATLAB,aes(x=-7,y=22, label = paste("sigma==", round(precision,2), sep = "")),  parse = TRUE,size=fontSz)
show(g)
```