### 2018-03-01 E1/doInferentialStats.Rmd
long to wide was a nightmare

---
title: "InspectHistogramsE1"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in parameter estimates

```{r}

estimates<- read.csv(file="Results/backwards2E1_paramEstimates.csv", header=TRUE, sep=",")
estimates$X<- NULL #Not sure where X comes from but it's just a row number
```

Exclude Ss who failed likelihood ratio test in all four cases.
Create new column nPassed for in how many conditions they passed (out of 4).

```{r}
library(dplyr)

estimates<- estimates %>% mutate(mixModelFitsBetter = pLRtest<.05)

condtnVariableNames <- c("subject","stream", "orientation")

e<- estimates %>%  group_by(subject) %>%
  summarise(nPassed = sum(mixModelFitsBetter)) 
estimates<- merge(e,estimates)

estimates<- estimates %>% mutate( exclude = (nPassed==0)  )
```

Assess the size of the effect to inform Bayesian test of next experiment (probably use raw units)

```{r, echo=FALSE, message=FALSE}

gg<-ggplot(estimates, aes(x=stream, y=efficacy)) + geom_point(color="grey") + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5)
show(gg)

library(tidyr)

data.table::dcast(df, orientation~stream,value.var="efficacy")

data.table::dcast(df, stream~orientation,value.var="efficacy")

temp<-estimates[,1:5]
ok<-data.table::dcast(temp, orientation+subject~stream,value.var="efficacy")
#Good except how do you preserve nPassed? Well, nPassed is the same for every factor combintation
data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="efficacy")
#Good so now try with two DVs. 
temp<-estimates[,1:6]
data.table::dcast(temp, orientation+subject+nPassed~stream,
                  value.var=c("efficacy","latency")) #doesn't work
data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="latency")

#The below works but drops many variables, so think about merging later
eff<- data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="efficacy")
lat<-data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var="latency") 
pre<-data.table::dcast(temp, orientation+subject+nPassed~stream,value.var="precision")

#So now scale up to full dataframe

```

```{r}

temp<-estimates[1:10,1:5]
temp$nPassed<-NULL
data.table::dcast(temp, subject+orientation~stream,value.var="efficacy")

temp %>%
  gather(key, value, -subject, -orientation, -nPassed, -stream) %>%
  unite(col, key, stream) %>%
  spread(col, value)

data.table::dcast(temp, subject+orientation+nPassed~stream,value.var="efficacy")

foo = data.frame(subject=c("AA","AA","AA","AA","AB","AB","AB","AB"),stream=c("Left","Left","Right","Right","Left","Left","Right","Right"),orientation=c("Canonical","Inverted","Canonical","Inverted","Canonical","Inverted","Canonical","Inverted"),efficacy=c(0.8,0.4,0.5,0.7,0.84,.37,.36,.69))
data.table::dcast(foo, subject+orientation~stream,value.var="efficacy")
foo.short<- foo[,1:4] #,c(2,3,4)]
f<-foo[1:4,2:4]
data.table::dcast(f, orientation~stream,value.var="efficacy")

tempp<-estimates[,4:6]
temppp<- tempp[,c(1,3)]

library(reshape2)

#data.table functions change things in place
require(data.table)
setDT(temp)
data.table::dcast(temp, subject+orientation+X+nPassed~stream,value.var="efficacy")

dcast(foo,group+num_users~times,value.var=c("action_rate","action_rate_c95"))

dcast(temp, subject+nPassed+X+orientation ~stream, value.var=efficacy)
?ddcast(tempp, 
temppp$id<- 1:nrow(temppp)
dcast(temppp, id~stream)
dcast(temppp, efficacy~stream+id)

wide <-  estimates %>%
  tidyr::spread(stream,efficacy)

df$

leftBias<- estimates %>% group_by(subject) %>%

deficitCanonical <- estimates %>% filter( stream=="Left" ) -

```

Do t tests.

Do Bayesian t-test. That is, take the likelihood of the left bias difference observed under the prior and divide it by the likelihood of the left bias difference observed under the null.

http://jeffrouder.blogspot.com.au/2016/01/what-priors-should-i-use-part-i.html
For the null model, the bias is just zero (delta function).

For our model, support is zero to infinity.

```{r, echo=FALSE, message=FALSE}
#want fig.height of 10 per subject

condtnVariableNames <- c("subject","stream", "orientation")

#Add R parameter estimates to dataframe
df<- merge(df,estimates)

```