### 2018-08-16 E2/analysis/importE2data.Rmd
after partially migrated from GoogleDrive

---
title: "importE2data"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

1/3 of trials have 1 target (numTargets==1)
targetSide ==0  means targets on both sides

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Modified from importE1data.R

```{r set up functions}

#Loads raw data from MATLAB files for second backwards paper second experiment
require(R.matlab)

mainPath<-".." # to main E2 folder

#Second experiment of second backwards-letters paper. random orientation each frame.
rawDataPath<- file.path(mainPath,
                            "Data/RawData/Data")
#raw data path containing .mat file for each subject

source(file.path("turnMATintoMeltedDataframe.R"))
#cat("Here are the files in the raw data directory")
#dir(rawDataPath)
```

```{r}
library(dplyr)
readInAllFiles<- function(rawDataPath) {

  files <- dir(path=rawDataPath,pattern='.mat')  #find all data files in this directory
  dfAll<-data.frame()
  for (i in 1:length(files)) { #read in each file
    fileThis<- file.path(rawDataPath,files[i])
    rawDataLoad=tryCatch(
      readMat(fileThis),
      error=function(e) {
        stop( paste0("ERROR reading the file ",fileThis," :",e) )
      } )
    apparentSubjectName <- strsplit(files[i],split="_")[[1]][1]
    subjectName<- rawDataLoad$participantID[[1]]
    if (apparentSubjectName != subjectName) {
      stop( paste0("WARNING apparentSubjectName",apparentSubjectName," from filename does not match subjectName in data structure",subjectName) )
    }
    else { cat(paste0(' Subject ',apparentSubjectName,' read in')) }
    rawDataLoad$file <- files[i]

    dfThis<- turnMATintoMeltedDataframe(rawDataLoad)

    tryCatch(
      dfAll<-rbind(dfAll,dfThis), #if fail to bind new with old,
      error=function(e) { #Give feedback about how the error happened
        cat(paste0("Tried to merge but error:",e) )
      } )
  }
  return(dfAll)
}
```

```{r read raw data}

E<- readInAllFiles(rawDataPath)
#TT<- as_tibble(E).  Seems to have no effect because weird hybrid type

#Calculate the serial position error
E$SPE<- E$respSP - E$targetSP

#give targetSide values that make  sense, like "dual", "left", and "right"
E<- E %>% mutate(targetSide = case_when(targetSide==0 ~ "both", 
                                       targetSide==1 ~"left", targetSide==2 ~"right",
                                       TRUE ~ "unexpected value"))
```

Next is very TIME CONSUMING.

Break out into target position -3, -2 etc. the targetStreamOrient

```{r veryTimeConsuming}
#Need to decompose into SPE=-3 orientation, SPE=-2 orientation, etc.
#Hm, can I use list apply to go through every row and pull out relevant bits?
#Very time-consuming!
E$orientMinus3<- -99; E$orientMinus2<- -99; E$orientMinus1<- -99; E$orient0<- -99; E$orientPlus1<- -99; E$orientPlus2<- -99;

for (i in 1:nrow(E)) {
  #row<- E[i,]
  #orientsThis <- row$targetStreamOrientations[[1]]
  #targetPos <- row$targetSP
  #If targetSP == NaN, it's a one-target trial and wrong side
  if (is.nan(E[i,]$targetSP)) {
    E[i,]$orientMinus3<- NaN
  } else {
    E[i,]$orientMinus3<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -3    ]
    E[i,]$orientMinus2<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -2    ]
    E[i,]$orientMinus1<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP -1    ]
    E[i,]$orient0<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP    ]
    E[i,]$orientPlus1<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +1    ]
    E[i,]$orientPlus2<- E[i,]$targetStreamOrientations[[1]][    E[i,]$targetSP +2    ]
  }
}

# More explicit way of doing it below but might take more time.
# for (i in 1:nrow(E)) {
#   row<- E[i,]
#   orientsThis <- row$targetStreamOrientations[[1]]
#   targetPos <- row$targetSP
#   #If targetSP == NaN, it's a one-target trial and wrong side
#   if (is.nan(targetPos)) {
#     E[i,]$orientMinus3<- NaN
#   }
#   E[i,]$orientMinus3<- orientsThis[targetPos-3]
#   #E[i,]$orientMinus3<- E[i,]$targetStreamOrientations[[1]][targetPos-3]
#   E[i,]$orientMinus2<- orientsThis[targetPos-2]
#   E[i,]$orientMinus1<- orientsThis[targetPos-1]
#   E[i,]$orient0<- orientsThis[targetPos] #should be same as $targetOrient
#   E[i,]$orientPlus1<- orientsThis[targetPos+1]
#   E[i,]$orientPlus2<- orientsThis[targetPos+2]
# }
```

Saving data.
```{r saveData}
backwards2_E2 <- E
pathToSave <- file.path( "..","Data")
saveRDS(backwards2_E2, file = "../Data/backwards2E1_rawDataFromMAT.rda", compress="bzip2")
```