### 2018-08-16 E2/analysis/inspectHists.Rmd

---
title: "inspect histograms"
author: "Alex Holcombe"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data

```{r load data, echo=FALSE, message=FALSE}

#Import raw data
dataPath<- file.path("../Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )

#tidy data
library(dplyr)
df<- data
#to work with dplyr, can't have array field like letterSeq
df$letterSeq<- NULL

```

Set up the experiment-specific parameters

```{r detailsNeeded}

numItemsInStream<- length( data$letterSeq[1,] )  
minSPE<- -17; maxSPE<- 17
annotateIt<-TRUE

numSs<-length(unique(df$subject))
cat( paste0('Total num Ss=',numSs) )
```

Plot sample histograms, make sure plotting working.

```{r}
dg<- df %>% filter(subject < "AC")

library(ggplot2)
g=ggplot(dg, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
show(g)

```

Plot all Ss
```{r, fig.height=200, fig.width=10}
g=ggplot(df, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times

show(g)
```

CY is bimodal but not outrageously so.

Subject AP looks to both have close to zero efficacy in some conditions and disregards instructions, reporting mostly after the cue.

BJ has essentially zero efficacy and may have tended to report letters from beginning of stream in one condition.

SsWhoLookWeird<-c("CY","AP","BJ")

We'll see whether the statistical comparison of mixture modeling to guessing distribution flags these as not being fit significantly better by mixture model.