### 2018-08-16 E2/analysis/mixModelE2data.Rmd

---
title: "mixModelE1data"
author: "Alex Holcombe"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,  comment = NA)
```

Load data

```{r load data, echo=FALSE, message=FALSE}

#Import data that was changed from MATLAB raw to R dataframe by importE2data.Rmd
dataPath<- file.path("..","Data")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )

#tidy data
library(dplyr)
df<- data
#to work with dplyr, can't have array field like letterSeq
df$letterSeq<- NULL

numItemsInStream<- length( data$letterSeq[1,] )  
minSPE<- -17; maxSPE<- 17

```

Why are there NaNs for targetSP?
```{r}
table(is.nan(df$targetSP))

```

Sanity-check some data by graphing.

Investigate whether pathological identicalness of inverted/canonical happens, that happened in E1 . Seems it doesn't

```{r}
dg<- df %>% filter(subject < "AC")

library(ggplot2)
g=ggplot(dg, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
show(g)
```

Try to analyze one bit of data.
WHY ARE THERE NANS IN TARGETSP?
```{r}
library(mixRSVP)
dg<-df %>% filter(subject=="AA")
analyzeOneCondition(dg,numItemsInStream,parameterBounds(),nReplicates=5)

analyzeOneConditionDF(df %>% filter(subject=="AA"),numItemsInStream,parameterBounds(),nReplicates=5)

```

Fit mixture model with R (if fitting not previously done and cached). 
```{r, echo=FALSE}

library(mixRSVP)

#Bizarrely, when knitting this often quits after only 6 parameter estimates. Don't know if there's some function fitting error catch problem

condtnVariableNames <- c("subject","stream", "orientation")

estimates<- df  %>% filter(subject=="AA") %>%
    group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
    do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )

head(estimates)
```

Save the param estimates to a file.
```{Save param estimates, echo=FALSE}

write.csv(estimates, file = "Results/backwards2E2_paramEstimates.csv")
#saveRDS(estimates, file = "Data/backwards2E1_paramEstimates.rda", compress="bzip2")

```
Done.