### 2018-08-19 E2/analysis/mixModelE2data.Rmd
progress

---
title: "mixModelE1data"
author: "Alex Holcombe"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,  comment = NA)
```

Load data

```{r load data, echo=FALSE, message=FALSE}

#Import data that was changed from MATLAB raw to R dataframe by importE2data.Rmd
dataPath<- file.path("..","Data")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )

#tidy data
library(dplyr)
#to work with dplyr, can't have array field like letterSeq, so before I fixed it in turnMatIntoMeltedDataframe, 
#I dropped the letterSeq thusly:
#df$letterSeq<- NULL

numItemsInStream<- length( data[1,"letterSeq"]$letterSeq$`1` )  
minSPE<- -17; maxSPE<- 17  #But it's 16 items in stream, so shouldn't it be -15, +15?

```

Hm, seems like funky StackOverflow way of turning matrix elements into list gave it this weird name "1" with backticks around it

data[1,"letterSeq"]$letterSeq$`1`

Make condition names more sensible - stream (Left/Right) and orientation (Canonical/Inverted)
```{r}
E2<- data
names(E2)[names(E2) == 'target'] <-'stream'
names(E2)[names(E2) == 'condition'] <-'orientation'
E2$stream[ E2$stream==1 ] <- "Left"
E2$stream[ E2$stream==2 ] <- "Right"
E2$targetSide[ E2$targetSide==0 ] <- "Both"
E2$targetSide[ E2$targetSide==1 ] <- "Left"
E2$targetSide[ E2$targetSide==2 ] <- "Right"

#mutate condition to Orientation
E2$orientation[ E2$orientation==1 ] <- "Canonical"
E2$orientation[ E2$orientation==2 ] <- "Inverted"

```

NaNs for targetSP indicate one-target trials, which are also indicated by targetSide==0
```{r}
table(E2$targetSP, E2$targetSide, dnn=c("targetSP","targetSide"), useNA="ifany")
```
Proportion of one-target trials seems to be 
`nrow(E2 %>% filter(targetSide!="Both")) / nrow(E2)`

Sanity-check some data by graphing.

Investigate whether pathological identicalness of inverted/canonical happens, that happened in E1 . Seems it doesn't

```{r}
dg<- E2 %>% filter(subject < "AC")

library(ggplot2)
g=ggplot(dg, aes(x=SPE)) + facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
show(g)
```

Try to analyze one bit of data.
```{r}
#devtools::install_github('alexholcombe/mixRSVP',build_vignettes=TRUE)
library(mixRSVP)
dg<-E2 %>% filter(subject=="AA")
dg<-dg %>% filter( !is.nan(targetSP) ) #excluding 1-target trials rows reporting stream on opposite side. Eventually, use those to evaluate swaps.

analyzeOneCondition(dg,numItemsInStream,parameterBounds(),nReplicates=2)

analyzeOneConditionDF(dg,numItemsInStream,parameterBounds(),nReplicates=2)

```

Analyse one subject, all conditions.
```{r, echo=FALSE}
#one<- dg %>% filter(targetSide != "Both")

condtnVariableNames <- c("subject","targetSide", "stream", "orientation")

estimatesAA<- dg  %>%  
    group_by_(.dots = condtnVariableNames) %>%  #.dots needed when you have a variable containing multiple factor names
    do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )
```

Fit mixture model (if fitting not previously done and cached). 
```{r, echo=FALSE}
#devtools::install_github('alexholcombe/mixRSVP',build_vignettes=TRUE)
library(mixRSVP)

#Bizarrely, when knitting this often quits after only 6 parameter estimates. Don't know if there's some function fitting error catch problem

estimates<- E2  %>%  filter( !is.na(targetSP) ) %>%
    group_by_at(.vars = condtnVariableNames) %>%  
    do(  analyzeOneConditionDF(.,numItemsInStream,parameterBounds(),nReplicates=5) )

head(estimates)
```

Calc numObservations for each condition
```{r }
#Calc numObservations for each condition. This is needed  for scaling the fine-grained Gaussian
#gaussianScaledforData needs to know.
dfGroups<-  E2  %>%  filter( !is.na(targetSP) ) %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
#add nPerCond back to parameter estimates
estimates<- merge(estimates,dfGroups)
```

Save the param estimates to a file.
```{r SaveParamEstimates, echo=FALSE}

resultsPath<- file.path("Results")

write.csv(estimates, file = file.path(resultsPath, "backwards2E2_paramEstimates.csv"))
#saveRDS(estimates, file = file.path(resultsPath, "Results/backwards2E1_paramEstimates.rda"), compress="bzip2")

```
Done.