### 2018-08-19 E2/analysis/doEverything.Rmd
working on oneTarget plotting

---
title: "inspect histograms"
author: "Alex Holcombe"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Optional validation (don't need to run this), loads raw data on its own and validates a few things

```{r validate data, echo=FALSE, message=FALSE}
source('validateE2data.Rmd')
```

Load data

```{r load data, echo=FALSE, message=FALSE}
source('importE2data.Rmd')
```

inspect Hists for weird data, non-performing Ss

```{r inspectHists}

source('inspectHists.Rmd')
```

Mixture model data

```{r}
source('mixModelE2data.Rmd')
```

Plot hists with psychophysical curves added,
check that and likelihood ratio test for whether to exclude, 
and infer
```{r}
source('inspectHistsWithFits.Rmd')
```