### 2018-08-23 E2/analysis/histsWithFitsOneTarget.Rmd
excluding

---
title: "one target"
author: "Alex Holcombe"
date: "2/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)

```

Load data and estimates

```{r load data, echo=FALSE, message=FALSE}

#Import raw data
dataPath<- file.path("../Data/")
#Experiment was administered by MATLAB
#.mat file been preprocessed into melted long dataframe by importE1data.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_rawDataFromMAT.rda") )

resultsPath<- file.path("Results")
estimates<- read.csv(file = file.path(resultsPath,"/backwards2E2_paramEstimates.csv"))

```

Fix the variable names if not done

```{r, echo=FALSE, message=FALSE}
#Below should now be done by import
if ('target' %in% names(data)) {
  names(data)[names(data) == 'target'] <-'stream'
  names(data)[names(data) == 'condition'] <-'orientation'
  data$stream[ data$stream==1 ] <- "Left"
  data$stream[ data$stream==2 ] <- "Right"
}
if (is.numeric(data$targetSide)) {
  data$targetSide[ data$targetSide==0 ] <- "Both"
  data$targetSide[ data$targetSide==1 ] <- "Left"
  data$targetSide[ data$targetSide==2 ] <- "Right"
}
if (is.numeric(data$orientation)) {
  #mutate condition to Orientation
  data$orientation[ data$orientation==1 ] <- "Canonical"
  data$orientation[ data$orientation==2 ] <- "Inverted"
}
```

Add the parameter estimates to the tibble

```{r, echo=FALSE, message=FALSE, fig.height=360, fig.width=10}
#want fig.height of 10 per subject

condtnVariableNames <- c("subject","stream", "orientation","targetSide")

#Add R parameter estimates to data tibble
#To merge them, unfortunately have to convert to dataframe first because dplyr has no binding of tibbles.
# https://stackoverflow.com/questions/45643886/how-to-combine-tibbles-with-common-but-non-identical-columns#45644319
#Not sure what is lost as a result.
dd<- as_data_frame(data)
es<- as_data_frame(estimates)
#Looks like each subject has only 8 efficacy values somehow, it's because there was an extra
#variable, X
es$X <- NULL
dat<- merge(dd,es)
dg<-dat %>% filter( !is.na(targetSP) ) #excluding 1-target trials rows reporting stream on opposite side. Eventually, use those to evaluate swaps.
```

Set up the experiment-specific parameters

```{r detailsNeeded}

numItemsInStream<- length( dg$letterSeq[[1]] )  
minSPE<- -17; maxSPE<- 17
annotateIt<-TRUE

numSs<-length(unique(dg$subject))
cat( paste0('Total num Ss=',numSs) )
```

Calc numObservations for each condition
```{r }
#Calc numObservations for each condition. This is needed  for scaling the fine-grained Gaussian
#gaussianScaledforData needs to know.
if (!("nPerCond" %in% colnames(dg))) {
  dfGroups<-  dg  %>%  filter( !is.na(targetSP) ) %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
  #add nPerCond back to parameter estimates
  estimates<- merge(estimates,dfGroups)
}
```

Calculate curves for parameters, to plot on histograms.
```{r, echo=TRUE, message=FALSE, fig.height=360, fig.width=10}
#devtools::install_github('alexholcombe/mixRSVP',build_vignettes=TRUE)
library(mixRSVP)
oneTarget<- dg %>% filter(targetSide!="Both")
curvesOneTarget<- dg %>% filter(targetSide!="Both") %>%
  group_by_at(.vars = condtnVariableNames) %>% 
  do(calc_curves_dataframe(.,minSPE,maxSPE,numItemsInStream))
#fails with on-missing arguments to min; returning Inf no non-missing arguments to max; returning -Inf 
#Error in createGuessingDistribution(minSPE, maxSPE, df$targetSP, numItemsInStream) :  minSPE must be less than or equal to maxSPE

#DEBUG
# dd<- dg %>% filter(subject=="AA", stream=="Left",targetSide=="Both",orientation=="Canonical")
# 
# createGuessingDistribution(minSPE,maxSPE,dd$targetSP,numItemsInStream)
# targetSP<- dd$targetSP
# maxTargetSP <- max(targetSP)
# minSPEthisData<- 1 - max(targetSP)
# maxSPEthisData<- numItemsInStream - min(targetSP)
# if (maxSPEthisData > maxSPE) stop("maxSPE must be at least",maxSPEthisData,"based on the values you passed me")
# if (minSPEthisData < minSPE) stop("minSPE must be no greater than",minSPEthisData,"based on the values you passed me")
# calc_curves_dataframe(dd,minSPE,maxSPE,numItemsInStream)

```

```{r }
# #Calc numObservations to each condition. This is needed only for scaling the fine-grained Gaussian
# #Calc the number of observations for each condition, because gaussianScaledforData needs to know.
# dfGroups<- dg %>% group_by_at(.vars = condtnVariableNames) %>% summarise(nPerCond = n())
# #add nPerCond back to parameter estimates
# estimates<- merge(estimates,dfGroups)

grain<-.05
gaussFine<- estimates %>%  filter(targetSide!="Both") %>% group_by_at(.vars = condtnVariableNames) %>% do(
  gaussian_scaled_from_df(.,minSPE,maxSPE,grain) )
```

Plot sample histograms, make sure plotting working.

```{r singleHistPlot}
library(ggplot2)

dAC<- dplyr::filter(oneTarget,subject=="AC")  
plotContinuousGaussian<- TRUE 
annotateIt<-TRUE
g<- plot_hist_with_fit(dAC,minSPE,maxSPE,d915$targetSP,numItemsInStream,
                        plotContinuousGaussian,annotateIt, FALSE)
g <- g + geom_vline(xintercept = 0)
g + annotate("text", x = 12, y = 25, label = "AC")
```

Show single plot with curves using scalable method to facet_grid.
In other words, make sure can plot without plot_hist_with_fit

```{r singleHistPlotScalable, echo=FALSE}

dAC<- dplyr::filter(oneTarget,subject=="AC")  
curvesThis <- curvesOneTarget %>% filter(subject=="AC")

g=ggplot(dAC, aes(x=SPE)) + 
        facet_grid(orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
g<-g +xlim(minSPE,maxSPE)
sz=.3
#show(g)
g<-g+ geom_point(data=curvesThis,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)

#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

#mixSig - whether mixture model statistically significantly better than guessing
curvesThis <- dplyr::mutate(curvesThis, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
g<- annotate_fit(g,curvesThis) #assumes curvesDf includes efficacy,latency,precision
#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
#identical commands are used. I haven't been able to work out why.
#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
show(g)
```

Plot all Ss
```{r, fig.height=200, fig.width=10}

g=ggplot(oneTarget, aes(x=SPE)) + 
        facet_grid(subject+orientation~stream) #,  scales="free_y")
g<-g+geom_histogram(binwidth=1,color="grey90") + xlim(minSPE,maxSPE)
g<-g+ geom_text(x=12, y= 33, aes(label = subject)) #inset subject name/number. Unfortunately it overwrites itself a million times
g<-g +theme_apa() #+theme(panel.grid.minor=element_blank(),panel.grid.major=element_blank())# hide all gridlines.
g<-g +xlim(minSPE,maxSPE)
sz=.3
#show(g)
g<-g+ geom_point(data=curvesOneTarget,aes(x=x,y=combinedFitFreq),color="chartreuse3",size=sz)
#g<-g+ geom_line(data=curves,aes(x=x,y=guessingFreq),color="yellow",size=sz)
#Discretized Gaussian
#g<-g+ geom_line(data=curves,aes(x=x,y=gaussianFreq),color="lightblue",size=sz)

#mixSig - whether mixture model statistically significantly better than guessing
curvesOneTarget <- dplyr::mutate(curvesOneTarget, mixSig = ifelse(pLRtest <= .05, TRUE, FALSE)) #annotate_fit uses this to color the p-value
g<- annotate_fit(g,curvesOneTarget) #assumes curvesDf includes efficacy,latency,precision
#Somehow the which mixSig (TRUE or FALSE) is red and which green is flipped relative to plot_hist_with_fit even though
#identical commands are used. I haven't been able to work out why.
#g<- g + scale_color_manual(values=c("red","forestgreen")) #already done in annotate_fit
show(g)

```

```{r investigateDifficultyCollapsing}
#see preservingNonNumericWhenSummarizing.R
tw<- as_tibble(twoTarget)
tw<- tw %>% select(-targetStreamOrientations,-warnings,-letterSeq)
ty<- tw %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), first(.))))

tt<- as_tibble(twoTarget) 
zb<- tt %>% group_by(subject,stream,orientation) %>% select(-targetStreamOrientations,-letterSeq) %>%
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

#The below works, but remember I've eliminated the list and string columns
#zz<- tw %>% group_by(subject,stream,orientation) %>%
#  summarise_all(function(x){ifelse(is.numeric(x),mean(x),last(x))})

tt<- as_tibble(twoTarget)
za<- tt %>% group_by(subject,stream,orientation) %>%
  summarise_all(function(x){ifelse(is.numeric(x),mean(x),last(x))})

tz<-tw
tz$targetSide <- as.factor(tz$targetSide)
tz %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), first(.))))

tw %>% group_by(subject,stream,orientation) %>% summarise_at("targetSide",first)

ty<- tw %>% group_by(subject,stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), -999.9)))
```

BE has e=.47 on left, .62 on right
BJ is just near chance and fails likelihood ratio test for 1-target, like did for 2-targets.
BO did badly for canonical on left, but passed lr test.
CF did badly for inverted on left, but passed lr test.
CL did pretty badly but got lucky, passed.
AP flunked one of the lr tests which is good because flunked some in two-target too.
So, lr test seems to be working well enough.

## Exclusions

Exclude by likelihood ratio test and efficacy.

For each subject, calculate how many two-target streams they failed. If 3 or 4 of 4, exclude (didn't think to mention 1-target condition in prereg).

Also check whether average efficacy is below 10% in both streams in either the upright or the reversed orientation.

```{r exclusions}

twoTarget<- dg %>% filter(targetSide=="Both")

AA<- twoTarget %>% filter(subject=="AA")
#Function for one subject, calculating how many stream conditions failed the test
calcExclude<- function(oneSub) {
  oneSub %>% group_by(stream,orientation)
}
#first, collapse over trials
#https://stackoverflow.com/questions/51978519/summarise-numeric-columns-return-last-value-of-non-numeric
tw<- twoTarget %>% group_by(subject,stream,orientation) %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

#for each subject, count pLRtest instances >.05 and efficay <.1
B<-tw %>% ungroup %>% group_by(subject) %>% summarise(LRtestFails = sum(pLRtest>.05), effFails = sum(efficacy<.1))                                                              
#B %>% print(n=Inf)
exclude <- B %>% filter(LRtestFails>=3, effFails >

# efficacy is below 10% in both streams in either the upright or the reversed orientation.
# To calculate that, need to group by subject and orientation
C<-tw %>% ungroup %>% group_by(subject,orientation) %>% summarise(bothEffsFailed = sum(efficacy<.1)==2)
#Find which subjects, if any, has a bothEffsFailed==TRUE
lowEff<- C %>% filter(bothEffsFailed==TRUE)
lowEff <- unique(lowEff$subject)

A<- tw %>% filter(subject=="AA")
B<-A %>% ungroup %>% summarise(LRtestFails = sum(pLRtest<.05), effFails = sum(efficacy<.1))

twoTarget %>% select(-targetStreamOrientations,-warnings,-letterSeq) %>% group_by(stream,orientation) %>% summarise_all(funs(if_else(is.numeric(.), mean(.), 3))

two<- twoTarget %>% group_by(stream,orientation) %>% summarise_all(funs(if_else(!is.list(.), mean(.), last(.))))

  summarise_all(funs(mean,sd))
A <- two %>% filter(subject=="AA")

curvesOneTarget<- twoTarget %>%
  group_by(subject) %>% 
  do(calc_curves_dataframe(.,minSPE,maxSPE,numItemsInStream))

```

Participant AI, BB suspicion of  attending to one side from twoTargets condition.