### 2018-08-29 E2/analysis/doEverything.Rmd
this is going well

---
title: "Do everything"
author: "Alex Holcombe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

The idea of this document is to have a single file that when knit does everything, [calling subsidiary Rmd documents](https://stackoverflow.com/questions/41962434/source-code-from-rmd-file-within-another-rmd)

But the problem is, the importE2data.Rmd throws a bizarre error when knit, but not when run!
Moreover, importE2data takes a very long time to run.
So in practice, the better thing to do is to run the below files one by one.

Optional validation (don't need to run this), loads raw data on its own and validates a few things

```{r validate, child = 'validateE2data.Rmd'}
```

Load data

```{r import, child = 'importE2data.Rmd'}
```

Inspect Hists for weird data, non-performing Ss.  This is prior to mixture modeling. Will do again after mixture modeling.

```{r inspectHists, child = 'inspectHists.Rmd'}
```

Mixture model data

```{r inspectHists, child = 'mixModelE2data.Rmd'}
```

Plot two-target hists with psychophysical curves added,
check that and likelihood ratio test for whether to exclude, 
```{r, child= 'histsWithFitsTwoTarget.Rmd'}
```

Plot one-target hists with psychophysical curves added,

```{r, child= 'histsWithFitsOneTarget.Rmd'}
```

Do exclusions, add column indicating whether subject excluded to dataframe, save that.
Should also add handedness quotient from handednessQuotientsP2E2 but there are so few left-handers it wouldn't make a difference. Need to pool across papers to test handedness issue.
```{r, child ='exclusions.Rmd')}
```

Look at incidence of location swaps.

What will I do about orientation? I can see how often they get orientation wrong, but that by itself is not so interesting.
I can look at the serial position correlogram of orientation and letter identity and I expect to see quite a strong correlation, contrary to what was found for two letters but something it would make sense to report at same time as Cheryl's finding.