### 2018-08-29 E2/analysis/dev/Chapter1.Rmd
right

---
title: "Chapter1"
author: "Alex Holcombe"
date: "8/28/2018"
output: html_document
---

Optional validation (don't have to run this to go to next step), loads raw data on its own and validates a few things

1/3 of trials have 1 target (numTargets==1)
targetSide ==0  means targets on both sides

Checking some stuff out

```{r set up functions}

#Loads raw data from MATLAB files for second backwards paper second experiment
require(R.matlab)

mainPath<-".." # to main E2 folder

#Second experiment of second backwards-letters paper. random orientation each frame.
rawDataPath<- file.path(mainPath,
                            "Data/RawData/Data")
#raw data path containing .mat file for each subject

source(file.path("turnMATintoMeltedDataframe.R"))
cat("Here are the files in the raw data directory")
dir(rawDataPath)
```

# Chapter 1

This is chapter 1.

```{r}
1
```