### 2018-08-30 E2/analysis/inferentialStats.Rmd
cool

---
title: "InferentialStats"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in parameter estimates

```{r}
dataPath <- file.path( "..","Data")
#Created by exclusions.Rmd
data<- readRDS( file.path(dataPath, "backwards2E2_dataWithEstimatesAndExcludeColumn.rda") ) 
data<- as_tibble(data)
```

For inferential stats, don't care about individual trials, so extract parameter estimates from full tibble.

```{r}

estimates<- data %>% select(-targetStreamOrientations,-letterSeq) %>% #can't include list columns
  group_by(subject,targetSide,orientation,stream) %>% #in other words, average over trials 
  summarise_all( function(x){ if(is.numeric(x)){ return(mean(x)) } else{ nth(x,-1) } })

library(readr)
resultsPath = "Results/"
write_tsv(estimates, file.path(resultsPath,"backwards2E2_paramEstimatesWithExcludeColumn.tsv"))

```

Plot data

```{r, echo=FALSE, message=FALSE}
require(ggplot2)
gg<-ggplot(estimates %>% filter(targetSide=="both"), 
           aes(x=stream, y=efficacy, color=orientation)) + geom_point() + stat_summary(fun.y="mean",geom="point",size=8, alpha=.5) + theme_apa() +ggtitle('nobody excluded')
show(gg)

hh<-ggplot(estimates %>% filter(targetSide=="both", exclude==FALSE), aes(x=stream, y=efficacy, color=orientation)) + 
  geom_point(position = position_dodge(width = 0.3)) + stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
  theme_apa() +ggtitle('after exclusions') + 
  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))

show(hh)

```

Graph the latency and sigma as well
BG, BJ, and AP have sigma near 4, clear outliers.

```{r, echo=FALSE, message=FALSE}
require(ggplot2)

hh<-ggplot(estimates %>% filter(targetSide=="both", exclude==FALSE), aes(x=stream, y=latency, color=orientation)) + 
  geom_point(position = position_dodge(width = 0.3)) +  stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
  theme_apa() +ggtitle('after exclusions') + 
  stat_summary(fun.data =mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))
show(hh)

ii<-ggplot(estimates %>% filter(targetSide=="both", exclude==FALSE), aes(x=stream, y=precision, color=orientation)) + 
  geom_point(position = position_dodge(width = 0.3)) +  stat_summary(fun.y="mean",geom="point",size=4, alpha=.5, position = position_dodge(width = 0.3)) + 
  theme_apa() +ggtitle('after exclusions') + 
  stat_summary(fun.data =mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = 0.3))
show(ii)

```

Convert from long to wide so that both left and right stream estimates are in the same row, while going from wide to long for efficacy,latency,precision so can use facet_grid with them.

```{r, echo=FALSE, message=FALSE}

library(data.table)
#The below works but drops many variables, so think about merging later
#I guess what comes after the "~" is what goes from long to wide
eff<- data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
                        value.var="efficacy")
eff$dv<-"efficacy"
lat<-data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
                       value.var="latency") 
lat$dv<-"latency"
pre<-data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
                       value.var="precision") 
pre$dv<-"precision"
all<-rbind(eff,lat,pre)
#I tried to do the above with a single command but makes an error even though StackOverflow example works
#https://stackoverflow.com/questions/37332076/from-long-to-wide-data-with-multiple-columns/37332345
# data.table::dcast(estimates, orientation+subject+nPassed~stream,value.var=c("efficacy","latency"))
head(all)
```

Calculate and plot bias. 
```{r , fig.height=20, fig.width=5}
all$leftBias<- all$Left - all$Right
hh<-ggplot(all %>% filter(targetSide=="both"), 
           aes(x=orientation, y=leftBias)) + facet_grid(dv~., scales="free") +
    stat_summary(fun.y="mean",geom="point",size=4, alpha=.5) + geom_point() +
  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.2) + theme_apa() +

show(hh)
```

Print effect size
```{r}
leftBiasesCanonical<- filter(all,dv=="efficacy" & orientation=="Canonical")$leftBias
leftBiasCanonical <- mean( leftBiasesCanonical )
leftBiasesInverted<- filter(all,dv=="efficacy" & orientation=="Inverted")$leftBias
leftBiasInverted<- mean(   leftBiasesInverted  )
leftBiasDiff<- leftBiasCanonical - leftBiasInverted
leftBias_sd <- sd(   leftBiasesCanonical - leftBiasesInverted )
```

The leftBiasDiff for this experiment is `r leftBiasDiff` or in Cohen's d `leftBiasDiff / leftBias_sd`, very healthy.

```{r, echo=FALSE}
E1n<- 16
#Raw effect size plus standard error and sd
P1E1canonical=.23;  P1E1c_se = .058; P1E1c_sd = P1E1c_se * sqrt(E1n)
P1E1reversed= .062;  P1E1r_se = .07; P1E1r_sd = P1E1r_se * sqrt(E1n)
#To calculate sd of the bias diff, consider that the variances sum,
#therefore first sum the variances then convert back to standard deviation
#Although only true for *independent* random variables and these unlikely to be totally independent
P1E1diff_sdCalculated = sqrt( (P1E1c_sd^2 + P1E1r_sd^2) )  #.257

E2n<- 24
P1E2canonical=.218; P1E2c_se =.034; P1E2c_sd= P1E2c_se * sqrt(E2n)
P1E2inverted=.017; P1E2i_se =.049; P1E2i_sd = P1E2i_se * sqrt(E2n)
P1E2diff_sdCalculated = sqrt( (P1E2c_sd^2 + P1E2i_sd^2) ) #.2066

P1E2vertUpright=.188; P1E2vu_se=.043; P1E2vu_sd = P1E2vu_se * sqrt(E2n)
P1E2vertInverted= -.07; P1E2vi_se=.049; P1E2vi_sd = P1E2vi_se * sqrt(E2n)
P1E2vdiff_sdCalculated = sqrt( (P1E2vi_sd^2 + P1E2vu_sd^2) )

P1E1leftBiasDiff<- P1E1canonical-P1E1reversed
P1E2leftBiasDiff<- P1E2canonical - P1E2inverted
P1E2vertBiasDiff<- P1E2vertUpright - P1E2vertInverted
#reported as  .258 +/- .047 in the paper

#The diff standard errors are also available from the paper directly. But still need to 
#convert to SD to get Cohen's d.

P1E1diff_se<-.07; P1E1diff_sd<- P1E1diff_se*sqrt(E2n) 
P1E2diff_se<-.038; P1E2diff_sd<- P1E2diff_se*sqrt(E2n) 
P1E2vdiff_se<- .047; P1E2vdiff_sd<- P1E2vdiff_se*sqrt(E2n)

```

What was it for the first paper? For E1,`r P1E1canonical`  Canonical and 
`r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
. For E2, bias difference very similar at `r P1E2leftBiasDiff`.
For E2 vertical arrangement, `r P1E2vertBiasDiff`

Do conventional stats.
```{r}

require(ez)

aa <- ezANOVA(data=filter(all,dv=="efficacy"), dv=leftBias, within=orientation, wid=subject)
cat("F=", aa$ANOVA$F, " p=", aa$ANOVA$p, "\n", sep="")

```

Do Bayesian t-test. That is, take the likelihood of the left bias difference observed under the prior and divide it by the likelihood of the left bias difference observed under the null.

Use the default Bayes test (but for E2 will use a custom prior)

```{r, echo=FALSE}
if(!require(BayesFactor)){install.packages('BayesFactor')}
library(BayesFactor)

leftBiasDifferences<- leftBiasesCanonical - leftBiasesInverted

ttestResults<- ttestBF( leftBiasDifferences )

```

The Bayes factor from a one-sample default Bayes t-test for the leftBiasDifferences is `r ttestResults@bayesFactor$bf`

The full results are 
```r 
print(ttestResults)
```

A custom prior would provide more power, as detailed in bayesianAnalysisForPrereg.Rmd in E2 folder

Do the correlogram and correlation