### 2018-08-31 E2/analysis/correlograms.Rmd
writing

---
title: "correlograms"
author: "Alex Holcombe"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Need to calculate the SPE of the orientation error. Unfortunately, it's ambiguous. Therefore, I will have multiple variables: orient0reported, orientPlus1reported, etc.   
And want to make a correlogram wherein

Read in data, parameter estimates

```{r}
dataPath <- file.path( "..","Data")
#Created by exclusions.Rmd
ff<- readRDS( file.path(dataPath, "backwards2E2_dataWithEstimatesAndExcludeColumn.rda") ) 
ff<- as_tibble(ff)
```

First look before collapsing wihtin subjects
```{r}
#I think I need to put the data in wide form so that it counts each combinatino of SPEleft and SPE right
#SPE<-data.table::dcast(ff %>% filter(exclude==FALSE), orientation+subject+targetSide~stream,
#                       value.var="SPE") 
SPE<-data.table::dcast(ff %>% filter(exclude==FALSE, numTargets==2), subject+trial~stream,
                       value.var="SPE") 
#Now get the counts so can plot heatmap
SPEcount<- SPE %>% count(Left,Right)
#plot that heatmap
correl<- ggplot(SPEcount, aes(x = Left, y = Right)) + geom_tile(aes(fill = n)) # +  facet_grid(separation~.)
correl

#Zero in on center because most large SPEs are guesses
#Plot with limits to zoom in
correl + xlim(-2,2) + ylim(-2,2)

```

What was it for the first paper? For E1,`r P1E1canonical`  Canonical and 
`r P1E1reversed` Mirror-reversed. Difference is `r P1E1canonical-P1E1reversed`
. For E2, bias difference very similar at `r P1E2leftBiasDiff`.
For E2 vertical arrangement, `r P1E2vertBiasDiff`

Probably want to calculate a correlation within each subject.

For correlogram

Could use below to plot average SPE for each condition, but should probably make it from scratch
```{r}
estimates <- read_tsv("backwards2E2_paramEstimatesWithExcludeColumn.tsv")
SPE<-data.table::dcast(estimates, orientation+subject+exclude+targetSide~stream,
                       value.var="SPE") 
```

The below is copied from code I wrote for Cheryl

#Calculate correlation
cor.test(df$responsePosRelativeleft, df$responsePosRelativeright,
         alternative = c("two.sided"),
         method = c("pearson"),
         exact = NULL, conf.level = 0.95, continuity = FALSE)

smallSPEs <- df %>% filter(abs(responsePosRelativeleft) < maxSPE ) %>% filter( abs(responsePosRelativeright) < maxSPE )

#ddply(smallSPEs, .(wordEcc), 
near<- smallSPEs %>% filter( wordEcc==1 )
far <- smallSPEs %>% filter( wordEcc > 1 )

cor.test(smallSPEs$responsePosRelativeleft, smallSPEs$responsePosRelativeright,
         alternative = c("two.sided"),
         method = c("pearson"),
         exact = NULL, conf.level = 0.95, continuity = FALSE)

nearCorr<- cor.test(near$responsePosRelativeleft, near$responsePosRelativeright,
                    alternative = c("two.sided"),
                    method = c("pearson"),
                    exact = NULL, conf.level = 0.95, continuity = FALSE)

farCorr<- cor.test(far$responsePosRelativeleft, far$responsePosRelativeright,
                   alternative = c("two.sided"),
                   method = c("pearson"),
                   exact = NULL, conf.level = 0.95, continuity = FALSE)

subjectCorr <- dlply(df,.(filename), function(x) cor.test(x$responsePosRelativeleft, x$responsePosRelativeright,
                                                          alternative = c("two.sided"),
                                                          method = c("pearson"),
                                                          exact = NULL, conf.level = 0.95, continuity = FALSE) )
farSubjectCorr <- dlply(far,.(filename), function(x) cor.test(x$responsePosRelativeleft, x$responsePosRelativeright,
                                                          alternative = c("two.sided"),
                                                          method = c("pearson"),
                                                          exact = NULL, conf.level = 0.95, continuity = FALSE) )
nearSubjectCorr <- dlply(near,.(filename), function(x) cor.test(x$responsePosRelativeleft, x$responsePosRelativeright,
                                                              alternative = c("two.sided"),
                                                              method = c("pearson"),
                                                              exact = NULL, conf.level = 0.95, continuity = FALSE) )

#nearCorrs <- c(.0711,.5278,.2991,.1807,.4680,.3630,-.0264,.3778,.4048,.1536,.1868,.0261,.1062,.1441,.0160)
#farCorrs <- c(-.2472,0,.2684,-.0428,.3736,-.0826,-.3119,-.0456,.1658,-.0603,-.2156,.0542,.1151,.1744,-.1492)
#Participants <- c(12,13,14,15,16,17,22,23,24,25,26,27,28,32,34)
#as.data.frame(allCorrData)
#allCorrData <- cbind(Participants,nearCorrs,farCorrs)
#t.test(nearCorrs,farCorrs, alternative = c("two.sided"), mu = 0, paired = FALSE, var.equal=FALSE, conf.level = 0.95)

#To-do
#Eliminate the five bad Ss [DONE!]
#Do the correlation analysis on each S individually, and then do paired t-tests on the subject means

#ADVANCED
#Address the argument that it wasn't correlated in the far just because of more guessing
#  - Do simulation of same correlation but with more guesses to make sure it is still significant despite the differing guessing rates
#Calculate for each subject and then do a t-test for near and far
#Analyse the frequent bigrams and actual words separately.