### 2018-09-03 manuscript/manuscript.Rmd
right

---
title             : "Backwards letters paper 2"
shorttitle        : "Title"

author: 
  - name          : "Alex O. Holcombe"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "School of Psychology, Brennan MacCallum Bldg (A19), The University of Sydney, Sydney, NSW 2006, Australia"
    email         : "my@email.com"
  - name          : "Christopher Bush"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "The University of Sydney"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line. This manuscript was written in RMarkdown using the *papaja* package [@R-papaja].

  Enter author note here.

abstract: |
  When one reads, early visual stages process the letters and words in parallel, but at a later stage, letters may need to be selected by attention to be fully processed. This transition to limited-capacity processing has recently been studied with a task requiring identification of simultaneously-presented letters embedded in a longer stream. Performance is markedly worse for identifying two letters rather than one. Although the task bears little resemblance to natural reading, implicit reading direction based on the letters' orientation nevertheless can determine which letter is poorly reported. This might be a result of word recognition processes rapidly and automatically extracting allocentric or object-relative reference frames. Alternatively, prioritisation based on implicit reading order (a left bias, for canonically-oriented letters) may not occur without a repeated regularity, such as the same orientation being used on multiple trials. In previous investigations, the trials were blocked, with all letters in the stream having the same orientation. Here, we first showed strong reading-order prioritization when trial order was random. Second, we found that when each frame within the stream had a random orientation, the bias was no longer affected by letter orientation, and defaulted to a left bias. Analysis of the errors supports the theory that the two streams of letters are sampled simultaneously, and a high-level process prioritizes one stream for consolidation into memory. 
  <!-- 250 word max for Memory & Cognition, https://www.springer.com/psychology/cognitive+psychology/journal/13421  -->

keywords          : "letters, words, attention, reading, allocentric"
wordcount         : "X"

bibliography      : ["referencesInfo/r-references.bib","referencesInfo/references.bib"]

floatsintext      : no
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

As you read these words, early visual stages process the letters and words in parallel. But at a later stage, letters may need to be selected by attention to be fully processed [@reichle_serial-attention_2011; @treisman_feature_1980]. This transition to limited-capacity processing has recently been studied with a task requiring identification of simultaneously-presented letters embedded in a longer stream. Performance is markedly worse for identifying two letters rather than one. Although the task bears little resemblance to natural reading, implicit reading direction based on the letters' orientation nevertheless can determine which letter is poorly reported. This might be a result of word recognition processes rapidly and automatically extracting allocentric or object-relative reference frames. Alternatively, prioritisation based on implicit reading order (a left bias, for canonically-oriented letters) may not occur without a repeated regularity, such as the same orientation being used on multiple trials. In previous investigations, the trials were blocked, with all letters in the stream having the same orientation. Here, we first showed strong reading-order prioritization when trial order was random. Second, we found that when each frame within the stream had a random orientation, the bias was no longer affected by letter orientation, and defaulted to a left bias. Analysis of the errors supports the theory that the two streams of letters are sampled simultaneously, and a high-level process prioritizes one stream for consolidation into memory.

The hypotheses, stopping rule, and other details of both experiments was preregistered (E1: https://osf.io/jdy8p/register/565fb3678c5e4a66b5582f67; E2: )

# Experiment 1

For the first experiment, 
 https://osf.io/jdy8p/register/565fb3678c5e4a66b5582f67

The experiment compares two within-subject levels of letter orientation (canonical or inverted), participants will experience both conditions within a single experimental session with the conditions presented in a mixed order. The method is almost exactly the same as that of Holcombe, Nguyen, & Goodbourn (in press). The only non-trivial difference is that in that experiment the two conditions were presented in separate blocks, whereas here they are mixed. This will allow us to determine whether simply seeing the backwards letters on an individual trial is sufficient to stop left-side prioritization, or whether having repeated successive trials of a particular condition is necessary. This speaks to how automatic the left-side prioritization observed in @holcombe_implied_2017 is.

If the left-side prioritisation is caused by implied reading direction then the left side bias will be stronger for letters presented in canonical orientation compared to inverted letters. Based on the findings of @holcombe_implied_2017, we predict that the left bias will be significantly stronger in canonical orientation than inverted.
We predict that latency and precision will not differ significantly between letter orientations

## Methods

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

Visual stimuli were generated using MATLAB R2012b with Psychtoolbox-3 (Brainard, 1997; Kleiner, Brainard, & Pelli, 2007; Pelli, 1997) on a MacBook Pro and displayed on a Mitsubishi DiamondView touchscreen LCD monitor with a resolution of 1,024 x 768 pixels and nominal refresh rate of 56 Hz (Experiment 2)CHRISCHECK. The code was adapted from the code used in a previous publication [@holcombe_implied_2017] that was written by [Patrick Goodbourn](https://orcid.org/0000-0001-7899-7355).

### Participants

Our previous study [@holcombe_implied_2017] of the effect of rotating the letters so they faced to the left yielded a large effect size of Cohen’s d = .94, which requires a sample size of only 14 to detect with 95% power. However, we expected that mixing the conditions within a block, as done here, might result in an effect size somewhat smaller, which led us to plan a sample size of 45 before exclusions (more details at the preregistration).

Forty-five participants were run (XXX females).

### Exclusions

As planned, six participants whose efficacy in either RSVP stream in any condition was lower than 10% (as indicated by the mixture modeling code) were removed from data analysis to reduce floor effects.

### Procedure

Preregistration: https://osf.io/r327j/
http://htmlpreview.github.io/?https://github.com/alexholcombe/backwardsLtrs2ndPaper/blob/master/E2/analysis/bayesianAnalysisForPrereg.html , which used our custom prior (details in statistical models section) and found that 75 participants would yield a decent Bayes Factor (about 10 or more) for the effect size we are expecting.

## Data analysis
The analysis code and results are all online (https://github.com/alexholcombe/backwardsLtrs2ndPaper). `r cite_r("referencesInfo/r-references.bib")` was used for the analyses.

### Mixture modeling
In a traditional analyses of performance in cued , the dependent variable would simply be accuracy at reporting the cued letter, for the left and for the right stream. However, that simple analysis can conceal differences in the dynamics of target selection from the streams [@vul_temporal_2008]. To address this, we developed mixture modeling [@goodbourn_pseudoextinction:_2015], which accommodates the possibility that in different conditions or with different streams, the latency of selection or temporal variability of selection may vary together with or separately from the rate of success of selecting a letter.

To perform mixture modeling, the data are coded as *serial position errors*. On each trial, for each stream, the participant chooses a letter from the choice array presented at the end of the trial. All the letters in the choice array were in fact presented, but they occurred in a random order. Thus, the letter presented can be expressed as a serial position error. The serial position error is 0 if the letter in the ring cue is reported. The serial position is -1 if the letter presented just before the cue is reported. It is +2 if the letter reported was that presented two after the cue And so on.

A histogram of the temporal errors forms a prominent hill whose peak is around the time of the cue (it almost always peaks at the time of the cue or one letter after). In addition, there is a "floor" to the histogram which formed by both very large and small errors, which appears to conform to the shape predicted by the participant making a random guess or not perceiving the target letter.

The mixture modeling code [@R-mixRSVP] fits the temporal error distribution with a combination of a Gaussian distribution (capturing the hill) and the guessing distribution. A maximum likelihood procedure finds the proportion of trials that can be attributed to the Gaussian, as opposed to the guessing distribution, and finds the best-fitting parameters ($\mu$, $\sigma$). Overall, then, we have the proportion of trials in which participants likely reported a letter around the time of the target, which we call *efficacy*, and the latency ($\mu$) and precision ($\sigma$).

# Results

The results of Experiment 1 were very similar to those of @holcombe_implied_2017.
<!-- (ref:my-figure-caption)
This is a figure caption for my figure.
fig.cap = "(ref:my-figure-caption)"
-->
<!-- fig.cap style fig captions example here: https://osf.io/ujwvr/?show=revision-->
```{r, fig.cap="Efficacy of ", echo=FALSE, message=FALSE}
require(readr); require(dplyr)
estimates <- read_tsv(file.path("..","E1","Results",'backwards2E1_paramEstimatesWithExcludeColumn.tsv'))
require(ggplot2)
require(papaja)
require(Hmisc) #for mean_cl_boot
dodg = .3
gg<-ggplot(estimates %>% filter(exclude==FALSE), 
           aes(x=orientation, y=efficacy, color=stream)) + 
          geom_point(position = position_dodge(width = dodg)) + 
  stat_summary(fun.y="mean",geom="point",size=8, alpha=.5, position = position_dodge(width = dodg)) +
  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
  theme_apa() 
show(gg)
```

```{r, ttestEffE1}
if(!require(BayesFactor)){install.packages('BayesFactor')}
library(BayesFactor)

leftCanon <- estimates %>% filter(exclude==FALSE,orientation=="Canonical",stream=="Left")
rightCanon <- estimates %>% filter(exclude==FALSE,orientation=="Canonical",stream=="Right")
leftInverted<-estimates %>% filter(exclude==FALSE,orientation=="Inverted",stream=="Left")
rightInverted<-estimates %>% filter(exclude==FALSE,orientation=="Inverted",stream=="Right")

bb<- BayesFactor::ttestBF(leftCanon$efficacy,rightCanon$efficacy)

```

As seen in Figure 1, when the text was canonically oriented, efficacy for the left stream (`r round(mean(leftCanon$efficacy),2)` ± `r round(sd(leftCanon$efficacy),2)` was much higher than for the right stream (`r round(mean(rightCanon$efficacy),2)` ± `r round(sd(rightCanon$efficacy),2)`), but when the text was inverted, efficacy for the left stream (`r round(mean(leftInverted$efficacy),2)` ± `r round(sd(leftInverted$efficacy),2)` was lower than for the right stream (`r round(mean(rightInverted$efficacy),2)` ± `r round(sd(rightInverted$efficacy),2)`)

efficacy was lower for the left stream () than for the right stream ().

"A total of 3 paired samples t-test will be conducted on left bias score (calculations below), latency and precision. All 3 t-tests will enter letter orientation (canonical vs inverted) as the independent variable."

Despite success at reporting the cued letter from the right stream occurring at a much lower rate

PLOT LEFT BIAS, will plot only these from now on.

```{r, fig.cap="Left bias", echo=FALSE, message=FALSE}

gg<-ggplot(estimates %>% filter(exclude==FALSE), 
           aes(x=orientation, y=efficacy, color=stream)) + 
          geom_point(position = position_dodge(width = dodg)) + 
  stat_summary(fun.y="mean",geom="point",size=8, alpha=.5, position = position_dodge(width = dodg)) +
  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
  theme_apa() 
show(gg)
```

apa_print()

As declared in the preregistration subsequent analyses will use "left bias" as the dependent variable, which is efficacy (or other dependent variable, such as latency) for the left stream minus that for the right stream.

A total of 3 paired samples t-test will be conducted on left bias score (calculations below), latency and precision. All 3 t-tests will enter letter orientation (canonical vs inverted) as the independent variable.

apa_print() and apa_table() facilitate reporting results of your analyses. Take a look at the .Rmd of the example manuscript in the folder example and the resulting .pdf.

Drop a supported analysis result, such as an htest- or lm-object, into apa_print() and receive a list of possible character strings that you can use to report the results of your analysis.

my_lm <- lm(Sepal.Width ~ Sepal.Length + Petal.Width + Petal.Length, data = iris)
apa_lm <- apa_print(my_lm)

# Discussion

# Experiment 2

## Method

### Exclusions

In the prereg, we wrote: "
Participants will be excluded if their average efficacy is below 10% in both streams in either the upright or the reversed orientation), or if the likelihood ratio test testing whether the data is fit better by the mixture model than by the guessing distribution fails (is not less than p<.05) for three or four of the four conditions. Because the Gaussian component of the mixture model is constrained to only allow latencies between -4 and 4, one effect of this is to eliminate participants who consistently do not report a letter near the time of the ring, instead for example reporting one of the last few items in the stream, which the data of past experiments indicates that rare participants do."

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup