### 2018-09-07 manuscript/manuscript.Rmd

---
title             : "Backwards letters paper 2"
shorttitle        : "Title"

author: 
  - name          : "Alex O. Holcombe"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "School of Psychology, Brennan MacCallum Bldg (A19), The University of Sydney, Sydney, NSW 2006, Australia"
    email         : "my@email.com"
  - name          : "Christopher Bush"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "The University of Sydney"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line. This manuscript was written in RMarkdown using the *papaja* package [@R-papaja].

  Enter author note here.

abstract: |
  When one reads, early visual stages process the letters and words in parallel, but at a later stage, letters may need to be selected by attention to be fully processed. This transition to limited-capacity processing has recently been studied with a task requiring identification of simultaneously-presented letters embedded in a longer stream. Performance is markedly worse for identifying two letters rather than one. Although the task bears little resemblance to natural reading, implicit reading direction based on the letters' orientation nevertheless can determine which letter is poorly reported. This might be a result of word recognition processes rapidly and automatically extracting allocentric or object-relative reference frames. Alternatively, prioritisation based on implicit reading order (a left bias, for canonically-oriented letters) may not occur without a repeated regularity, such as the same orientation being used on multiple trials. In previous investigations, the trials were blocked, with all letters in the stream having the same orientation. Here, we first showed strong reading-order prioritization when trial order was random. Second, we found that when each frame within the stream had a random orientation, the bias was no longer affected by letter orientation, and defaulted to a left bias. Analysis of the errors supports the theory that the two streams of letters are sampled simultaneously, and a high-level process prioritizes one stream for consolidation into memory. 
  <!-- 250 word max for Memory & Cognition, https://www.springer.com/psychology/cognitive+psychology/journal/13421  -->

keywords          : "letters, words, attention, reading, allocentric"
wordcount         : "X"

bibliography      : ["referencesInfo/r-references.bib","referencesInfo/references.bib"]

floatsintext      : no
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

As you read these words, early visual stages process the letters and words in parallel. But at a later stage, letters may need to be selected by attention to be fully processed [@reichle_serial-attention_2011; @treisman_feature_1980]. This transition to limited-capacity processing has recently been studied with a task requiring identification of simultaneously-presented letters embedded in a longer stream. Performance is markedly worse for identifying two letters rather than one. Although the task bears little resemblance to natural reading, implicit reading direction based on the letters' orientation nevertheless can determine which letter is poorly reported. This might be a result of word recognition processes rapidly and automatically extracting allocentric or object-relative reference frames. Alternatively, prioritisation based on implicit reading order (a left bias, for canonically-oriented letters) may not occur without a repeated regularity, such as the same orientation being used on multiple trials. In previous investigations, the trials were blocked, with all letters in the stream having the same orientation. Here, we first showed strong reading-order prioritization when trial order was random. Second, we found that when each frame within the stream had a random orientation, the bias was no longer affected by letter orientation, and defaulted to a left bias. Analysis of the errors supports the theory that the two streams of letters are sampled simultaneously, and a high-level process prioritizes one stream for consolidation into memory.

## Prioritization and coordinate frames

What causes the letter on one side to be accurately identified more often than the letter on the other side? An answer to this should have two parts. One part should be an account of which processing stages the letter on one side gains an advantage, as well as the nature of those processing stages. The answer must also explain what causes the particular side that is prioritized to be the side that is prioritized. The present experiments were designed to provide insight into the second question.
One possibility is that the prioritization emerges after repeated trials of a particular orientation. This is the "gradual emergence" possibility.

could in principle reflect the slow emergence of a cognitive strategy, albeit one that does not

the left side, say, to be prioritized rather than the right side.

prioritization to occ

A mental coordinate system must be used to allocate resources differentially to two locations, as a location can only be defined relative to a coordinate frame. To favor the first letter as defined by implicit reading direction, the mind must recognize the orientation of the letters and determine which of the two letters is on the left (for the canonical orientation) or which on the right (for the mirror-reversed or inverted condition). Studies of brain-injured patients have revealed aspects of how such relative spatial loca- tions are represented.
Among patients with parietal neglect, some neglect the left visual field, some neglect the left half of each word, and some neglect the first part of each word, even if the letters are mirror- reversed so that the first part of the word is to the right (Haywood & Coltheart, 2001; Hillis et al., 2005). The reference frame thought to underpin the last phenomenon is sometimes referred to as object-centered or word-centered (Caramazza & Hillis, 1990). Conceivably, it could be critical to the implicit reading-direction bias documented in the present experiments: If the two letters acted like a single word, the word-centered frame would represent the location of the first letter that can then be prioritized for processing. Although some scheme for calculating the first letter is clearly needed, we deem it unlikely that it is the same as that used to represent a word. We give our reasons here.
The present experiments used widely separated letters, which did not appear to form a single word or even a group. Moreover, the present finding of a very large advantage for the first letter relative to the last letter is not a typical characteristic of individual- word processing. Although the literature on individual-word rec- ognition does show that the first letters of a word are more informative than the last, the effect is not as large as the left advantage in our dual-RSVP paradigm (e.g., Blais et al., 2009). The RSVP literature does not seem to note a particularly large advantage either. In pilot experiments with a single RSVP stream of four-letter words (rather than two RSVP streams of individual letters), we have not observed a large advantage for the first two letters. Thus, it appears that the bias toward the first letter in our experiments does not reflect a word-centered processing frame and may be unrelated to the processes that yield word-centered neglect and normal word identification.
Although two letters within a word may be processed in parallel (Ans, Carbonnel, & Valdois, 1998; Blais et al., 2009; Coltheart, Rastle, Perry, Langdon, & Ziegler, 2001; Davis, 2010; Perry, Ziegler, & Zorzi, 2007; cf. Whitney & Cornelissen, 2008), some evidence indicates that widely spaced letters are identified serially (Cohen, Dehaene, Vinckier, Jobert, & Montavont, 2008; Risko, Lanthier, & Besner, 2011). This serial processing may be a char- acteristic of the extraction of spatial relationships between arbi- trary objects as well as letters. In particular, attentional shifts between objects may be critical to extracting the spatial relation- ship between them (Franconeri, Scimeca, Roth, Helseth, & Kahn, 2012; Holcombe, Linares, & Vaziri-Pashkam, 2011). Our laresults suggest that the two target letters are sampled in parallel, but a subsequent stage may process them serially. Reading expe- rience may have led to this learned visual routine (Ullman, 1984) of beginning with the first letter according to letter orientation. Training monkeys on spatial-relationship judgments can result in parietal neurons tuned to relative location for the trained stimuli (Olson, 2003). Such neurons may be involved in coding the relative location of the letters, in combination with task-related (reading) top-down signals (Filimon, 2015).

## The present studies

The hypotheses, stopping rule, and other details of both experiments was preregistered (E1: https://osf.io/jdy8p/register/565fb3678c5e4a66b5582f67; E2: )

# Experiment 1

For the first experiment, 
 https://osf.io/jdy8p/register/565fb3678c5e4a66b5582f67

The experiment compares two within-subject levels of letter orientation (canonical or inverted), participants will experience both conditions within a single experimental session with the conditions presented in a mixed order. The method is almost exactly the same as that of [@holcombe_implied_2017]. The only non-trivial difference is that in that experiment the two conditions were presented in separate blocks, whereas here they are mixed. This will allow us to determine whether simply seeing the backwards letters on an individual trial is sufficient to stop left-side prioritization, or whether having repeated successive trials of a particular condition is necessary. This speaks to how automatic the left-side prioritization observed in @holcombe_implied_2017 is.

If the left-side prioritisation is caused by implied reading direction then the left side bias will be stronger for letters presented in canonical orientation compared to inverted letters. Based on the findings of @holcombe_implied_2017, we predict that the left bias will be significantly stronger in canonical orientation than inverted.
We predict that latency and precision will not differ significantly between letter orientations

## Methods

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

Visual stimuli were generated using MATLAB R2012b with Psychtoolbox-3 (Brainard, 1997; Kleiner, Brainard, & Pelli, 2007; Pelli, 1997) on a MacBook Pro and displayed on a Mitsubishi DiamondView touchscreen LCD monitor with a resolution of 1,024 x 768 pixels and nominal refresh rate of 56 Hz (Experiment 2)CHRISCHECK. The code was adapted from the code used in a previous publication [@holcombe_implied_2017] that was written by [Patrick Goodbourn](https://orcid.org/0000-0001-7899-7355).

### Participants

Our previous study [@holcombe_implied_2017] of the effect of rotating the letters so they faced to the left yielded a large effect size of Cohen’s d = .94, which requires a sample size of only 14 to detect with 95% power. However, we expected that mixing the conditions within a block, as done here, might result in an effect size somewhat smaller, which led us to plan a sample size of 45 before exclusions (more details at the preregistration).

Forty-five participants were run (XXX females).

### Exclusions

As planned, six participants whose efficacy in either RSVP stream in any condition was lower than 10% (as indicated by the mixture modeling code) were removed from data analysis to reduce floor effects.

### Procedure

The letters in both streams were presented at a rate of 10 per second (66 ms presentation and 33 ms interstimulus interval). The two target letters, which were simultaneous, were surrounded by white cue rings of .1 deg thickness and 5 deg diameter. The screen was viewed from a distance of approximately 57 cm.

## Data analysis
The analysis code and results are all online (https://github.com/alexholcombe/backwardsLtrs2ndPaper). `r cite_r("referencesInfo/r-references.bib")` was used for the analyses.

### Mixture modeling
In a traditional analyses of performance in cued , the dependent variable would simply be accuracy at reporting the cued letter, for the left and for the right stream. However, that simple analysis can conceal differences in the dynamics of target selection from the streams [@vul_temporal_2008]. To address this, we developed mixture modeling [@goodbourn_pseudoextinction2015], which accommodates the possibility that in different conditions or with different streams, the latency of selection or temporal variability of selection may vary together with or separately from the rate of success of selecting a letter.

To perform mixture modeling, the data are coded as *serial position errors*. On each trial, for each stream, the participant chooses a letter from the choice array presented at the end of the trial. All the letters in the choice array were in fact presented, but they occurred in a random order. Thus, the letter presented can be expressed as a serial position error. The serial position error is 0 if the letter in the ring cue is reported. The serial position is -1 if the letter presented just before the cue is reported. It is +2 if the letter reported was that presented two after the cue And so on.

A histogram of the temporal errors forms a prominent hill whose peak is around the time of the cue (it almost always peaks at the time of the cue or one letter after). In addition, there is a "floor" to the histogram which formed by both very large and small errors, which appears to conform to the shape predicted by the participant making a random guess or not perceiving the target letter.

The mixture modeling code [@R-mixRSVP] fits the temporal error distribution with a combination of a Gaussian distribution (capturing the hill) and the guessing distribution. A maximum likelihood procedure finds the proportion of trials that can be attributed to the Gaussian, as opposed to the guessing distribution, and finds the best-fitting parameters ($\mu$, $\sigma$). Overall, then, we have the proportion of trials in which participants likely reported a letter around the time of the target, which we call *efficacy*, and the latency ($\mu$) and precision ($\sigma$).

# Results

The results of Experiment 1 were very similar to those of @holcombe_implied_2017.
<!-- (ref:my-figure-caption)
This is a figure caption for my figure.
fig.cap = "(ref:my-figure-caption)"
-->
<!-- fig.cap style fig captions example here: https://osf.io/ujwvr/?show=revision-->

(ref:effE1) Efficacy of .

```{r, fig.cap='(ref:effE1)', echo=FALSE, message=FALSE}
require(readr); require(dplyr)
estimates <- read_tsv(file.path("..","E1","Results",'backwards2E1_paramEstimatesWithExcludeColumn.tsv'))
require(ggplot2)
require(papaja)
require(Hmisc) #for mean_cl_boot
dodg = .3
gg<-ggplot(estimates %>% filter(exclude==FALSE), 
           aes(x=orientation, y=efficacy, color=stream)) + 
          geom_point(position = position_dodge(width = dodg)) + 
  stat_summary(fun.y="mean",geom="point",size=8, alpha=.5, position = position_dodge(width = dodg)) +
  stat_summary(fun.data = mean_cl_boot, geom="errorbar", width=0.4, position = position_dodge(width = dodg)) +
  theme_apa() 
show(gg)
```

```{r, ttestEffE1}
if(!require(BayesFactor)){install.packages('BayesFactor')}
library(BayesFactor)

leftCanon <- estimates %>% filter(exclude==FALSE,orientation=="Canonical",stream=="Left")
rightCanon <- estimates %>% filter(exclude==FALSE,orientation=="Canonical",stream=="Right")
leftInverted<-estimates %>% filter(exclude==FALSE,orientation=="Inverted",stream=="Left")
rightInverted<-estimates %>% filter(exclude==FALSE,orientation=="Inverted",stream=="Right")

bb<- BayesFactor::ttestBF(leftCanon$efficacy-rightCanon$efficacy,
                          leftInverted$efficacy-rightInverted$efficacy)
latexReport<- (apa_print(bb)$statistic)[[1]]
```

As seen in Figure 1, when the text was canonically oriented, efficacy for the left stream (`r round(mean(leftCanon$efficacy),2)` ± `r round(sd(leftCanon$efficacy),2)`) was markedly higher than for the right stream (`r round(mean(rightCanon$efficacy),2)` ± `r round(sd(rightCanon$efficacy),2)`), but when the text was inverted, efficacy for the left stream (`r round(mean(leftInverted$efficacy),2)` ± `r round(sd(leftInverted$efficacy),2)`) was lower than for the right stream (`r round(mean(rightInverted$efficacy),2)` ± `r round(sd(rightInverted$efficacy),2)`).

For our primary statistical analyses (as preregistered), we use "left bias" as the dependent variable. The evidence here is very strong, `r latexReport`, that the left bias is greater in the canonical condition (`r mean(leftCanon$efficacy - rightCanon$efficacy)`  ± `r sd(leftCanon$efficacy - rightCanon$efficacy)`) than in the inverted condition (`r mean(leftInverted$efficacy - rightInverted$efficacy)`  ± `r sd(leftInverted$efficacy - rightInverted$efficacy)`).

```{r latencyPrecision}
bflLR<- ttestBF(leftCanon$latency,rightCanon$latency)
latexReportBflLR<- (apa_print(bflLR)$statistic)[[1]]
bflLRinvert<- ttestBF(leftInverted$latency,rightInverted$latency)
latexReportBflLRinvert<- (apa_print(bflLRinvert)$statistic)[[1]]

bfl<- ttestBF(leftCanon$latency - rightCanon$latency, leftInverted$latency - rightInverted$latency)
latexReportBfl<- (apa_print(bfl)$statistic)[[1]]
bfp<- ttestBF(leftCanon$precision - rightCanon$precision, leftInverted$precision - rightInverted$precision)
latexReportBfp<- (apa_print(bfp)$statistic)[[1]]

```

The latency and precision were similar in all four conditions, such that the Bayesian t-test supported the null hypothesis of no difference between left and right in both conditions (canonical: `r latexReportBflLR`, inverted: `r latexReportBflLRinvert`). The latency values are:
canonical, left = `r round(mean(leftCanon$latency)*100)` ms, right = `r round(mean(rightCanon$latency)*100)` ms; inverted, left = `r round(mean(leftInverted$latency)*100)` ms, right = `r round(mean(rightInverted$latency)*100)` ms, and the precision values are: canonical, left = `r round(mean(leftCanon$precision)*100)` ms, right = `r round(mean(rightCanon$precision)*100)`; inverted, left = `r round(mean(leftInverted$precision)*100)`, right = `r round(mean(rightInverted$precision)*100)`. Thus, there was no support for the hypothesis of a difference between the left biases for the canonical and the inverted conditions, for latency (`r latexReportBfl`) or for precision (`r latexReportBfp`.

ANALYZE THE EFFECT OF RESPONSE ORDER

## No effect of previous trials

Many perceptual phenomena are subject to serial dependencies. To investigate if that is the case here, we assessed whether the left bias on a trial is affected only by the orientation of the letters on that trial, or is instead also affected by the orientation of the letters on previou trials. Mixture modeling could not be used to calculate a left bias for an individual trial, so instead of efficacy we used "approximately correct", which was also used for certain purposes in previous papers [e.g., @holcombe_implied_2017]. "Approximately correct" is defined as a serial position error of -1, 0, or 1. For an individual trial, then, left bias was this approximately correct value for the left stream minus approximately correct for the right stream.
For each participant, this trial-by-trial left bias was correlated with the orientation of the current and four previous trials. This is plotted in Figure \@ref(corrE1f) .

(ref:corrE1) Correlation of left bias with this trial's orientation and with previous trial orientations.

\@ref(corrE1f)

```{r corrE1f, fig.cap='(ref:corrE1)'}
require(knitr)
knitr::include_graphics(file.path("..","E1","Results","Plots","autocorrE1.png"))
```

```{r readStreakBF, echo=FALSE, message=FALSE}
 filePath<- file.path("..","E1","Results","streakBayesfactor.txt")
 con = file(filePath, "r")
 line = readLines(con, n = 1)
```

It is possible that the build-up of the bias is nonlinear in that the effect of a streak of trials with the same orientation is greater than the sum of the effect of the individual previous trials. We therefore compared the left bias after a streak of four trials with the same orientation to the left bias in all the trials that were not preceded by such a streak. However, there was no evidence for an effect, Bayesian t-test `r line`.

# Discussion

Experiment 1 replicated the usual [@goodbourn_pseudoextinction2015; @holcombe_implied_2017; @ransley_reading_2018] dissociation between the efficacy and the dynamics of selection as indexed by latency and precision. That is, despite spatial biases in efficacy, there was no spatial bias in the latency and precision. This suggests that the reason for the poor performance on the right in the canonical condition, and on the left in the inverted condition, was not due to a delay in attentional selection or failure to allocate attention on one side. It may instead be due to a post-selection processing failure. Perhaps a post-selection process has a severe capacity limitation, resulting in strong competition between the two letters.  
The novelty here was that the particular letter orientation (canonical or inverted) was chosen randomly on each trial. In previous work, trials with the same orientation were blocked, and it was unknown whether this was necessary. It might have been the case that multiple successive trials with the same orientation are required before the reading direction bias was set up in the mind.
In Experiment 2 we further probed the necessary conditions for the reading direction bias.

# Experiment 2

LOOK UP IRINA'S AND CORBALLIS' PAPERS ON ORIENTAITON, SEE IF THEY SAY ANYTHING ABOUT REFERENCE FRAMES.
One

Kim in her thesis says "spatial bias on early versus late trials on each block) showed little effect,"
COULD OPTIONALLY INCLUDE AN EXPERIMENT 3, JUST REPORTING THE PSYC1 CANONICAL VERSUS INVERTED TO SHOW IT DOESN'T NEED AN RSVP STREAM

## Method

Only differences from Experiment 1 are described below.

### Procedure

The letters in both streams were presented at a slightly slower rate than in Experiment 1, at 8 per second (80 ms presentation and 40 ms interstimulus interval).

Preregistration: https://osf.io/r327j/
http://htmlpreview.github.io/?https://github.com/alexholcombe/backwardsLtrs2ndPaper/blob/master/E2/analysis/bayesianAnalysisForPrereg.html , which used our custom prior (details in statistical models section) and found that 75 participants would yield a decent Bayes Factor (about 10 or more) for the effect size we are expecting.

### Exclusions

In the prereg, we wrote: "
Participants will be excluded if their average efficacy is below 10% in both streams in either the upright or the reversed orientation), or if the likelihood ratio test testing whether the data is fit better by the mixture model than by the guessing distribution fails (is not less than p<.05) for three or four of the four conditions. Because the Gaussian component of the mixture model is constrained to only allow latencies between -4 and 4, one effect of this is to eliminate participants who consistently do not report a letter near the time of the ring, instead for example reporting one of the last few items in the stream, which the data of past experiments indicates that rare participants do."

## Results

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup